FIXME

### 2013-02-10 stackExchange/writeups/iccm-2013/writeup.tex
foo
Former-commit-id: 14a45947fc03e3818491a9788bd8510b271c1d60 [formerly 4dd0d2a0b772728f982bffbbf758f7754ecb1f82]
Former-commit-id: 723742e46f278a8c71b1fe0c1e88b0c015f4401e

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act.pdf}}
  \caption{Example Post}
  \label{fig:examplePost}
\end{figure}

@@

A query was then used to build the post word, tag word co-occurrence matrix.
For a given word found in the body or title of a post, the query counts the number of times that that word appears with each tag.
Each of those counts is recorded as the word co-occurrence for that word and each associated tag.
This process is done across all words found in any post.
The result is a comma-separated value (csv) file of context word, tag, count for each combination.

@@

Model tag predictions are also based on the frequency that each tag has been used.
More often used tags more likely to have higher activation.
A SQL query was used to build a similar csv for tag occurrences.
This csv for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.

@@

\begin{table}[!ht]
  \begin{center} 
    \caption{Sample table title.} 
    \label{sample-table} 
    \vskip 0.12in
    \begin{tabular}{ll} 
      \hline
      Common Name &  Equation \\
      \hline
      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} + O$ \\
      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)}$ \\
      & 			$= log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$ \\
      Attentional Weight	& $W_{j} = WE_{j}$ \\
      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
      Max Entropy & 		$H_{max} = maxH$ \\
      \hline
    \end{tabular} 
  \end{center} 
\end{table}

@@

Contextual activation ($S_{ji}$) can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
For example, if the words in the title contained only 'PHP', 'PHP5', and 'PHPUnit', then the number of $_{j}$'s is three.
And for the 'PHP' tag, those three words are often in the title of PHP-tagged posts.
So the sum of the three $S_{ji}$'s for the tag PHP will be high, meaning that the odds adjustment for PHP given the current context will be high.
Since PHP already has a high base-level activation (frequently used tag), the model should produce a high total activation for the tag PHP, given the context.

@@

Attentional weighting ($W_{j}$) can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
For the StackOverflow model, the attentional weighting was left unbound, and allowed to be calibrated by the logistic regression model.
This was done for two main reasons.
First, the researcher was curious to see what value the logistic regression model would give for $W$.
Second, a non-ACT-R standard measure of attentional weighting was used ($E_{j}$), so it seemed appropriate to allow $W$ to vary alongside exploring this new weighting measure.

@@

R was chosen as the implementation language for two main reasons.
First, it has sparse matrix support and vectorization routines, so that the model could scale to millions of $S_{jis}$'s.
Second, it has excellent data visualization and analysis capabilities. 
This way statistical data analysis and visualization could be done within the same R process that was running the model.

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{foo.pdf}}
  \caption{
    Logistic regression of tag used on post (binary) regressed on four components of model tag activation.
  }
  \label{fig:logReg}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act.pdf}}
  \caption{
    Model's prediction of top-ten tags for post in Figure \ref{fig:examplePost}.
    The title is the title text of the post.
    The attentional weight (entropy) for each word in the title is included below each word.
    The user-chosen tags are included below the attentional weights.
    The stacked bar graphs are the total activation for the top-ten predicted model tags for the post.
    Each bar graph is partitioned into the 3 model components that vary between tags within a post (not the post offset).
}
  \label{fig:modelPost}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{ROC.pdf}}
  \caption{
    Model performance across the calibration and test dataset, as well as when specific model components are removed.
    The vertical dotted line represents the point where the model is asked to tag on average one tag per post.
  }
  \label{fig:ROC}
\end{figure}

>>>>

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act-cropped.pdf}}
  \caption{Example Post}
  \label{fig:examplePost}
\end{figure}

@@

A query was then used to build the post word, tag word co-occurrence matrix.
For a given word found in the body or title of a post, the query counts the number of times that that word appears with each tag.
Each of those counts is recorded as the word co-occurrence for that word and each associated tag.
This process is done across all words found in any post.
The result is a comma-separated value (csv) file of context word, tag, count for each of the \num{39223968} unique context word, tag combinations.

@@

Model tag predictions are also based on the frequency that each tag has been used.
More often used tags more likely to have higher activation.
A SQL query was used to build a similar csv of tag use counts for the \num{26176} unique tags.
This csv for example would show that C\# is the most often used tag, where about 11\% of posts were tagged with C\#.

@@

\begin{table}[!ht]
  \begin{center} 
    \caption{Sample table title.} 
    \label{sample-table} 
    \vskip 0.12in
    \begin{tabular}{ll} 
      \hline
      Common Name &  Equation \\
      \hline
      Activation & 		$A_{i} = B_{i} + \sum_{j\in T}^{ } W_{j} S_{ji} + \sum_{j\in B}^{ } W_{j} S_{ji} - O$ \\
      Base Level & 		$B_{i} = log \frac{p_{i}}{1-p_{i}}$ \\
      Strength Assoc. &		$S_{ji} = log \frac{p(i|j)}{p(i))} = log \frac{p(j,i)}{p(j)p(i)}$ \\
      & 			$= log \frac{NN(j,i)}{N_{Row}(j)N_{Col}(i)}$ \\
      Attentional Weight	& $W_{j} = WE_{j}$ \\
      Scaled Entropy & 		$E_{j} = 1-\frac{H_{j}}{H_{max}}$ \\
      Entropy & 		$H_{j} = -\sum_{i=1}^{N}p(i|j)log\left (  p(i|j) \right )$ \\
      Max Entropy & 		$H_{max} = maxH$ \\
      Offset & 			$O = \frac{1}{5}\sum_{i\in top 5}^{ } A_{i}$ \\
      \hline
    \end{tabular} 
  \end{center} 
\end{table}

@@

Contextual activation can be thought of as the amount of odds adjustment to base levels, given the current context $_{j}$.
Contextual activation has two components: strength of association ($s_{sji}$) and attentional weight ($w_{j}$).
Strength of association measures the relatedness of a contextual element $_{i}$ (e.g., the word PHP in the title of a post) with a tag $_{j}$ (e.g., the tag PHP5).

@@

The second component of contextual activation, attentional weighting ($W_{j}$), can be thought of as the amount of attentional resources that are dedicated to contextual element $_{j}$.
Total attentional weighting ($W$) is bound, most often set to 1, and reflects the fact that our attentional resources are limited.
For the StackOverflow model, the attentional weighting was left unbound, and allowed to be calibrated by the logistic regression model.
This was done for two main reasons.
First, the researcher was curious to see what value the logistic regression model would give for $W$.
Second, a non-ACT-R standard measure of attentional weighting was used ($E_{j}$), so it seemed appropriate to allow $W$ to vary alongside exploring this new weighting measure.

@@

\subsubsection{Offset}

An offset measurement (mean of the top five tags for the post) was used to equalize the top activated tags for each post.
This was done primarily to improve model fit.
To see why this is the case, suppose the following scenario:
The offset measurement is not used, and the model is asked to provide 10 tags across 10 posts.
The top ten tags for one of these posts have activations that are higher than activations for all other tags in the other nine posts.
The model would then select the ten tags from the single post as the most likely tags.
But this doesn't make intuitive sense, because there is a limit to the number of tags that people tend to use for each post (average number of tags per post is ~ 3).
So the offset measure aims to place all of these 10 posts on the same level, so that asking the model to choose 10 tags from 10 posts will select roughly a single tag from each post.

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{foo-cropped.pdf}}
  \caption{
    Logistic regression of tag used on post (binary) regressed on four components of model tag activation.
  }
  \label{fig:logReg}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{visPost-4148010-act-cropped.pdf}}
  \caption{
    Model's prediction of top-ten tags for post in Figure \ref{fig:examplePost}.
    The title is the title text of the post.
    The attentional weight (entropy) for each word in the title is included below each word.
    The user-chosen tags are included below the attentional weights.
    The stacked bar graphs are the total activation for the top-ten predicted model tags for the post.
    Each bar graph is partitioned into the 3 model components that vary between tags within a post (not the post offset).
}
  \label{fig:modelPost}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \resizebox{\linewidth}{!}{\includegraphics{ROC-cropped.pdf}}
  \caption{
    Model performance across the calibration and test dataset, as well as when specific model components are removed.
    The vertical dotted line represents the point where the model is asked to tag on average one tag per post.
  }
  \label{fig:ROC}
\end{figure}
